---
title: "Runboard"
output: 
  flexdashboard::flex_dashboard:
    css: styles.css
    theme: united
    orientation: rows
    social: menu
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(cycleRtools)
library(dplyr)
library(leaflet)
library(tidyr)
library(dygraphs)
library(xts)
library(stringr)
library(elevatr)
library(rgdal)
library(ggplot2)

# helpers
source("helpers/data.R")
```


Sidebar {.sidebar}
==================

### Open activity:

```{r}
fileInput('fitFile', label='Only .fit files are supported')
```


### Heart Rate Zones

```{r}
selectInput('hrmethod', 'HR based on', c('Lactate Threshold', 'Maximum HR'))
sliderInput('hrthresh', 'HR in bpm', 170, min=150, max=220)
```

### Elevation Correction

If your device does not have a barometric sensor, try to correct GPS data with a [external radar data](https://cran.r-project.org/web/packages/elevatr/vignettes/introduction_to_elevatr.html).

```{r}
checkboxInput('fixElevation', label='Fix Elevation', value=FALSE)
```

```{r}
numericInput('currentLocation', '', value=1)
dateInput('rangeStart', '', NULL)
dateInput('rangeStop', '', NULL)

```

```{r}

# takes forever - calls java *cough*
rawdata <- reactive({
  req(input$fitFile)
  elevationCache(NULL) # drop the cache
  read_fit(input$fitFile$datapath, format=F)
})

# data cleaning and fixing
fitdata <- reactive({
  df <- rawdata() %>%
    tbl_df %>%
    run_rename_select %>%
    run_fix_units %>%
    run_fix_elevation(skip=!input$fixElevation, cache=elevationCache()) %>%
    run_add_gain %>%
    run_add_pace
  if (input$fixElevation && is.null(elevationCache())) {
    elevationCache(df$altitude)
  }
  df
})

# limit data to range from dygraph
rangedata <- reactive({
  fitdata() %>%
    run_filter_fitrange(start=input$rangeStart, stop=input$rangeStop) %>%
    run_add_hrzone(breaks=hrBreaks(), labels=hrLabels())
})

mapdata <- reactive({
  tres <- fitdata() %>%
            mutate(td=timestamp - lag(timestamp)) %>%
            select('td') %>%
            drop_na() %>% 
            summarise(mean(td))
  rangedata() %>%
    filter(!(row_number() %% max(tres, 5))) # take everyt 5th for plotting
})

# depends on maximum hr in range
# used to restrict the extent of HR-Zone-plot
hrBreaks <- reactive({
  thresh <- input$hrthresh
  # https://www.researchgate.net/figure/The-5-Zone-Intensity-Scale-Used-by-the-Norwegian-Olympic-Federation-and-the-3-Zone-and_tbl1_259652575
  breaks <- thresh * c(0.72, 0.82, 0.87, 0.92, 1)
  if (str_detect(input$hrmethod, 'Lactate')) {
    breaks <- breaks / 0.92
  }
  breaks <- round(breaks)
  c(min(breaks[1]-1, min(fitdata()$hr)), breaks)
})

# statics
elevationCache <- reactiveVal(NULL)
hrLabels <- reactiveVal(c('Z1', 'Z2', 'Z3', 'Z4', 'Z5'))
hrColors <- c('#c4ced4', '#b0dcf7', '#b9dfa0', '#fbd89e', '#faa8b1')
hlCallbackJs <- 'function(event, x, points, row, seriesName){
  Shiny.setInputValue("currentLocation", row);
  for (div of document.getElementsByClassName("dygraphs")) {
    var obj = window.HTMLWidgets.getInstance(div);
    obj.dygraph.setSelection(row);
  }
}'
```

Summary
=======

Row
---

### Pace {.value-box}

```{r}
avgPace <- reactive({
  data <- rangedata()
  d <- ( tail(data$distance, n=1) - head(data$distance, n=1) ) / 1000
  t <- as.numeric(
    difftime(tail(data$timestamp, n=1), data$timestamp[1], unit='mins')
  )
  pace <- t/d
  pre <- floor(pace)
  list(string=sprintf('%d:%02d min/km', pre, floor(60 * (pace - pre))),
       numeric=pace)
})

renderValueBox({
  value <- avgPace()
  valueBox(value = value$string,
           icon = "fa-tachometer-alt",
           color = 'success')
           #color = if (value$numeric < 6) 'success' else 'primary')
})
```

### Distance {.value-box}

```{r}
distance <- reactive({
  data <- rangedata()
  d <- ( tail(data$distance, n=1) - head(data$distance, n=1) ) / 1000
  list(string=sprintf('%0.1f km', d), numeric=d)
})

renderValueBox({
  value = distance()
  valueBox(
    value = value$string,
    icon = "fa-road",
    color = 'success'
  )
    #color = if (value$numeric > 10) 'success' else 'primary')
})
```


### Elevation Gain (possibly imprecise) {.value-box}

```{r}
elGain <- reactive({
  gains <- rangedata()$gain
  gain <- floor(sum( gains[gains>0] ))
  list(string=sprintf('%d m', gain), numeric=gain)
})

renderValueBox({
  value <- elGain()
  valueBox(
    value = value$string,
    icon = "fa-mountain",
    color = 'success'
  )
})
```

### Average HR {.value-box}

```{r}
avgHR <- reactive({
  avghr <- round(mean(rangedata()$hr))
  list(string=sprintf('%d bpm', avghr), numeric=avghr)
})

renderValueBox({
  value <- avgHR()
  valueBox(
    value = value$string,
    icon = "fa-heartbeat",
    color = 'danger'
  )
})
```

Row
---

### Map {data-width=50%}

```{r}

output$map <- renderLeaflet({
  map <- leaflet() %>% 
    setView(lat=0, lng=0, zoom=2) %>%
    addTiles()
  map
})

# clear, when we load new data
observe({
  data <- rawdata() %>% drop_na()
  proxy <- leafletProxy('map')
  proxy %>%
    clearMarkers()
})

# remove current marker when range and/or input changes
observe({
  data <- mapdata() %>% select(c('timestamp', 'lat', 'long')) %>% drop_na()
  current <- fitdata() %>% slice(input$currentLocation)
  proxy <- leafletProxy('map')
  proxy %>%
    clearGroup('current') %>%
    addMarkers(lng=current$long, current$lat, group='current')
})

# draw current range
observe({
  data <- mapdata() %>% select(c('timestamp', 'lat', 'long', 'hrzone')) %>% drop_na()
  proxy <- leafletProxy('map')
  proxy %>%
    clearMarkers() %>%
    addCircleMarkers(lng=data[data$hrzone==hrLabels()[1], ]$long,
                     lat=data[data$hrzone==hrLabels()[1], ]$lat,
                     opacity=0.1, color='black', radius=8, fillColor=hrColors[1], fillOpacity=0.5) %>%
    addCircleMarkers(lng=data[data$hrzone==hrLabels()[2], ]$long,
                     lat=data[data$hrzone==hrLabels()[2], ]$lat,
                     opacity=0.1, color='black', radius=8, fillColor=hrColors[2], fillOpacity=0.5) %>%
    addCircleMarkers(lng=data[data$hrzone==hrLabels()[3], ]$long,
                     lat=data[data$hrzone==hrLabels()[3], ]$lat,
                     opacity=0.1, color='black', radius=8, fillColor=hrColors[3], fillOpacity=0.5) %>%
    addCircleMarkers(lng=data[data$hrzone==hrLabels()[4], ]$long,
                     lat=data[data$hrzone==hrLabels()[4], ]$lat,
                     opacity=0.1, color='black', radius=8, fillColor=hrColors[4], fillOpacity=0.5) %>%
    addCircleMarkers(lng=data[data$hrzone==hrLabels()[5], ]$long,
                     lat=data[data$hrzone==hrLabels()[5], ]$lat,
                     opacity=0.1, color='black', radius=8, fillColor=hrColors[5], fillOpacity=0.5) %>%
    fitBounds(min(data$long), min(data$lat), max(data$long), max(data$lat))
    #addPolylines(data=data, lat=~lat, lng=~long) %>%
})

observe({
  a <- input$rangeStart
  b <- input$rangeStop
})

leafletOutput('map')
```

### Charts {data-width=50%}

```{r}
fillCol(
  flex = c(5,7),
 
  renderDygraph({
    hrB <- hrBreaks()
    data <- fitdata()
    tickerJs <- 'function(min, max, pixels, opts, dygraph, vals){return([])}'
    
    dygraph(xts(data[c('hr')], order.by=data$timestamp),
            group='global') %>%
      dyOptions(colors=c('black'), fillAlpha=0.1, fillGraph=TRUE) %>%
      dyCallbacks(highlightCallback=hlCallbackJs) %>%
      dyAxis("x", drawGrid=FALSE) %>%
      dyAxis('y', valueRange=c(hrB[1], hrB[6] + 1), ticker=tickerJs) %>%
      dyLimit(hrB[1], hrB[1], labelLoc="left", color="black") %>%
      dyLimit(hrB[2], hrB[2], labelLoc="left", color="black") %>%
      dyLimit(hrB[3], hrB[3], labelLoc="left", color="black") %>%
      dyLimit(hrB[4], hrB[4], labelLoc="left", color="black") %>%
      dyLimit(hrB[5], hrB[5], labelLoc="left", color="black") %>%
      dyShading(hrB[1], hrB[2], color=hrColors[1], axis="y") %>%
      dyShading(hrB[2], hrB[3], color=hrColors[2], axis="y") %>%
      dyShading(hrB[3], hrB[4], color=hrColors[3], axis="y") %>%
      dyShading(hrB[4], hrB[5], color=hrColors[4], axis="y") %>%
      dyShading(hrB[5], hrB[6], color=hrColors[5], axis="y")
  }),
  
  renderDygraph({
    drawCallbackJs<- 'function(g, i){
      Shiny.setInputValue("rangeStart", Dygraph.dateString_(g.xAxisRange()[0], "UTC"))
      Shiny.setInputValue("rangeStop", Dygraph.dateString_(g.xAxisRange()[1], "UTC"))
    }'
    
    data <- fitdata()
    dygraph(xts(data[c('pace', 'altitude')],
                order.by=data$timestamp), group='global') %>%
      dyCallbacks(highlightCallback=hlCallbackJs,
                  drawCallback=drawCallbackJs) %>%
      dySeries("altitude", axis = 'y2') %>%
      dyRangeSelector() %>%
      dyAxis("y", valueRange=c(0, max(data$pace))) %>%
      dyAxis("y2", valueRange=c(min(data$altitude),
                                min(data$altitude) +
                        (max(data$altitude) - min(data$altitude)) *2)) %>%
      dyAxis("x", drawGrid = FALSE)
  }) 
) 
```

Analysis
========

Row {data-height=80}
---------------------

### min HR

```{r}
sliderInput('minHR', value=140, min=140, max=200,
            'Only take values into consideration that are above this heart rate.')
```

### HR group size
```{r}
sliderInput('hrGroupSize', value=1, min=1, max=5, step=1,
            'Try to smooth the index by grouping neighbouring heart rate values.')
```

### max gradient
```{r}
sliderInput('maxGain', value=0.05, min=0, max=1, step=0.005,
            'Remove all data pairs where the absolute estimated altitude gradient is smaller than this value (prefer flat segments).')
```

### min group size
```{r}
sliderInput('minGroup', value=15, min=1, max=100,
            'Ignore data pairs wich occured fewer than this value')
```


Row {.tabset}
---

```{r}
analysisdata <- reactive({
  rangedata() %>%
    select(c('gain', 'pace', 'hr')) %>%
    filter(hr >= input$minHR) %>%
    filter(abs(gain) <= input$maxGain) %>%
    mutate(hrg=floor(hr/input$hrGroupSize)*input$hrGroupSize) %>%
    group_by(hrg) %>%
    filter(n()>input$minGroup) %>%
    mutate(med=median(pace), avg=mean(pace),
           lwr=median(pace)-sd(pace), upr=median(pace)+sd(pace)) %>%
    filter((abs(pace- median(pace)) <= 2 * sd(pace)))
})
```

### HR-Pace Index

```{r}

renderPlot({
  data <- analysisdata()

  ggplot(data, aes(hrg, pace)) +
    scale_x_continuous(name='Heart Rate', breaks=unique(data$hrg)) +
    scale_y_continuous(trans = "reverse") +
    geom_point() +
    geom_smooth(method='loess', n=20)

})
```

### Gradient Histogram

```{r}
renderPlot({
  data <- analysisdata()

  ggplot(data, aes(abs(gain))) +
    geom_histogram()
})
```
